---
title: "Wilcoxon_test"
author: "Yutian "
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r install libraries, message=FALSE,results = 'hide',include=FALSE}
list.of.packages <- c("ggplot2", "dplyr","data.table","ggpubr","mice","missForest","Amelia","ggstatsplot","plotly","cowplot","gapminder","PMCMRplus")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages,repos = 'http://cran.us.r-project.org')
```



```{r load libraries, warning=FALSE, message=FALSE, results = 'hide',include=FALSE}
library(dplyr)
library(ggplot2)
library(data.table)
library(ggpubr)
library(mice)
library(missForest)
library(Amelia)
library(ggstatsplot)
library(plotly)
library(cowplot)
library(ggstatsplot)
```



From the density plot in the home page, it can be seen that `avg_glucose_level` is not normally distribution, therefore, Wilcoxon test will be performed to find out whether the difference in two groups is significant.

The Wilcoxon test, is a nonparametric statistical test that compares two paired groups. The tests essentially calculate the difference between sets of pairs and analyze these differences to establish if they are statistically significantly different from one another.

- H0: There is no statistically significant difference between people who suffered a stroke and people who did not suffer a stroke in this sample data

- H1: There is a statistically significant difference between people who suffered a stroke and people who did not suffer a stroke in this sample data

If the p-value is less than 0.05, we reject the null hypothesis that there's no difference between the means and conclude that a significant difference does exist

```{r,include=FALSE}
stroke <- read.csv("stroke_cleaned.csv")
```

```{r}
test <- wilcox.test(stroke$avg_glucose_level ~stroke$stroke_t)
test
```

The p-value (p-value < 0.05) indicates that we can reject the null hypothesis, and conclude that at the 5% significance level average glucose level are significantly different between stroke patients and subjects without stroke. 



```{r s1,include=FALSE}
# plot with statistical results
s1 <- ggbetweenstats( # independent samples
  data = stroke,
  x = stroke_t,
  y = avg_glucose_level,
  plot.type = "box", # for boxplot
  type = "Non-parametric", # for wilcoxon
  centrality.plotting = FALSE, # remove median,
  title = "Comparison of average glucose level in stroke & Non-stroke patients",
  pairwise.display = FALSE,
  pairwise.comparisons = FALSE,
)
```


```{r g2,include=FALSE,warning=FALSE}
## for reproducibility
set.seed(123)


## select part of the dataset and use it for plotting
g2<- grouped_ggbetweenstats(
    ## arguments relevant for ggbetweenstats
   data = stroke,
  x = stroke_t,
  y = avg_glucose_level,
    grouping.var = age_level,
    xlab = "Stroke Status",
    ylab = "Avereage Glucose Level",
    pairwise.display = "significant", ## display only significant pairwise comparisons
    p.adjust.method = "fdr", ## adjust p-values for multiple tests using this method
    # ggtheme = ggthemes::theme_tufte(),
    package = "ggsci",
   type = "np",
    palette = "default_jco",
    outlier.tagging = TRUE,
    outlier.label =  age_level,
    ## arguments relevant for combine_plots
    annotation.args = list(title = "Age_level"),
    plotgrid.args = list(nrow = 3)
  )

```


```{r g3,include=FALSE,warning=FALSE}
## for reproducibility
set.seed(123)
## select part of the dataset and use it for plotting
g3<-
  grouped_ggbetweenstats(
    ## arguments relevant for ggbetweenstats
   data = stroke,
  x = stroke_t,
  y = avg_glucose_level,
    grouping.var = obesity_level,
    xlab = "Stroke Status",
    ylab = "Avereage Glucose Level",
    pairwise.display = "significant", ## display only significant pairwise comparisons
    p.adjust.method = "fdr", ## adjust p-values for multiple tests using this method
    # ggtheme = ggthemes::theme_tufte(),
    package = "ggsci",
    palette = "default_jco",
    type = "np",
    outlier.tagging = TRUE,
    outlier.label =  obesity_level,
    ## arguments relevant for combine_plots
    annotation.args = list(title = "Obesity_level"),
    plotgrid.args = list(nrow = 3)
  )
```


# Wilcoxon test


Null and alternative hypothesis of the Wilcoxon test are as follows:


## Supplementary Figures {.tabset}

### Supplementary Figure 1


```{r, warning=FALSE,echo=FALSE}
plotly::ggplotly(s1, width = 480, height = 480) 
```

It is an interactive plot, although this box plot displays a higher median average glucose level value in stroke patients, the scatters show that the average glucose level of stroke patients is concentrated above 200(mg/dl) or below 110(mg/dl), with few in the middle.

### Supplementary Figure 2

```{r, warning=FALSE,echo=FALSE}
g2
```

This box violin plot divided objects by age level and conducted Significant tests at each age category. There are no newborn stroke patients and only 2 child stroke patients in this subject which are not sufficient to conduct the significant test. And in adults, the p-value is 0.01, so we can not reject the null hypothesis, while the p-value in the elderly is less than 0.05, therefore we can conclude that average glucose level is significantly different between old stroke patients and old non-stroke patients at a 5% significance level.


### Supplementary Figure 3

```{r, warning=FALSE,echo=FALSE}
g3
```

This box violin plot divided objects by obesity level and conducted Significant tests at each BMI category. In the obese category, the p-value is less than 0.05, therefore we can conclude that the average glucose level is significantly different between obese stroke patients and old non-stroke patients at a 5% significance level.






